A Discontinuous Petrov-Galerkin method for convection-diffusion problems

### 2013-07-30 docs/Jesse/Dissertation/NSNotes.tex

We briefly review the compressible Navier-Stokes equations, given in Section~\secref{sec:compNS}, and formulate DPG for the nonlinear system. 
\begin{itemize}
\item \textbf{Conservation equations}
\begin{align*}
\div \vecttwo{\rho u_1 }{\rho u_2} &= 0\\
\div \vecttwo{\rho u_1^2+p }{\rho u_1 u_2} &= \div\left(\vec{\sigma_{i1}}\right)\\
\div \vecttwo{\rho u_1 u_2}{\rho u_2^2+p } &= \div\left(\vec{\sigma_{i2}}\right)\\
\div\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} &= \div\left[\boldsymbol\sigma U + \vec{q}\right],
\end{align*}
where $\boldsymbol \sigma$ is the stress tensor whose $ij$th term is $\sigma_{ij}$.

@@

We represent $\boldsymbol\sigma$ using the Newtonian fluid law
\[
\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}
\]
where $\mu$ is viscosity and $\lambda$ is bulk viscosity. 
%Following elasticity, a general formula for the stress tensor is 
%\[
%\sigma_{ij} = 2\mu \epsilon_{ij} + \lambda \epsilon_{kk} \delta_{ij}
%\]
We can invert the stress tensor under isotropic and plane strain assumptions to get
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
We also have
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \grad  U - \boldsymbol \omega
\]
where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
\[
\boldsymbol \omega = \frac{1}{2}\left(\grad  U - \grad ^T  U\right).
\]
Thus our final form is
\begin{align*}
\grad  U - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
Notice that $\boldsymbol \omega$ is implicitly defined to be the symmetric part of $\grad u$ by taking the symmetric part of the above equation.

@@

To summarize, our system of equations in the classical variables is now
\begin{align*}
\div \vecttwo{\rho u }{\rho v} &= 0\\
\div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol\sigma \mathbf{u} + \vec{q}\right) &=0\\
\frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I &= \grad \mathbf{u} - \Reyn \, {\boldsymbol \omega}\\
\frac{1}{\kappa}\vec{q} &= \grad T
\end{align*}
We strongly enforce symmetry of the stress tensor $\boldsymbol \sigma$ by setting $\sigma_{21} = \sigma_{12}$. Additionally, we have scaled the antisymmetric tensor $\boldsymbol \omega$ by the Reynolds number to ensure that $\boldsymbol \omega = O(1)$ for all ranges of $\Reyn$.

@@

@@

The Navier-Stokes conservation laws can be written as 
\begin{align*}
\div \vecttwo{\rho u }{\rho v} &= 0\\
\div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0
\end{align*}
or generally, 
\[
\div (F_i(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \sigma) = 0, \qquad i = 1,\ldots, 4
\]
The variational form restricted to a single element gives
\[
\langle \widehat{F}_i\cdot n, v\rangle - \int_K  (F(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \sigma)) \cdot \grad v_i = 0 , \qquad i = 1,\ldots, 4
\]
and the variational form over the entire domain is given by summing up the element-wise contributions.

The presence of terms such as $\boldsymbol \sigma_i \cdot \boldsymbol u$ means that we will need to linearize in the stress variables $\sigma_ij$ in addition to our Eulerian quantities. Since fluxes and traces are linear, we do not need to linearize them. Instead, fluxes $\widehat{F}_{i,n}$ and traces $\widehat{u},\widehat{v},\widehat{T}$ will represent normal traces and traces of the accumulated nonlinear solution. The linearized variational formulation is thus
\begin{align*}
\langle \widehat{F}_i\cdot n, v\rangle &- \int_K  \left(F_{i,\boldsymbol U}(\boldsymbol U)\cdot \Delta \boldsymbol U -G_{i,\boldsymbol U}(\boldsymbol U, \boldsymbol \sigma)\cdot \Delta \boldsymbol U - G_{i,\boldsymbol \sigma}(\boldsymbol U, \boldsymbol \sigma)\cdot \Delta \boldsymbol \sigma \right)\cdot \grad v_i \\
&= \int_K  \left(F_i(\boldsymbol U)-G_i(\boldsymbol U)\right) \cdot \grad v_i \\
\qquad i &= 1,\ldots, 4
\end{align*}
where $F^i_{j,\boldsymbol U}$, $G^i_{j,\boldsymbol U}$, and $G^i_{j,\boldsymbol \sigma}$ are the Eulerian and two viscous Jacobians (linearized w.r.t.\ the Eulerian/viscous variables), respectively.
%where
%\begin{align*}
%F^1_{1,\boldsymbol U} &= \{u,\rho ,0,0\} \\
%F^2_{1,\boldsymbol U} &= \{v,0,\rho ,0\} \\
%F^1_{2,\boldsymbol U} &=\left\{c_v T (\gamma -1)+u^2,2 u \rho ,0,c_v (\gamma -1) \rho \right\}\\
%F^2_{2,\boldsymbol U} &=\{u v,v \rho ,u \rho ,0\}\\
%F^1_{3,\boldsymbol U} &=\{u v,v \rho ,u \rho ,0\}\\
%F^2_{3,\boldsymbol U} &=\left\{c_v T (\gamma -1)+v^2,0,2 v \rho ,c_v (\gamma -1) \rho \right\}\\
%F^1_{4,\boldsymbol U} &=\left\{\frac{1}{2} u \left(2 c_v T (2 \gamma -1)+u^2+v^2\right),\frac{1}{2} \rho  \left(2 c_v T (2 \gamma -1)+3 u^2+v^2\right),u v \rho ,c_v u (2 \gamma -1) \rho \right\}\\
%F^2_{4,\boldsymbol U} &=\left\{\frac{1}{2} v \left(2 c_v T (2 \gamma -1)+u^2+v^2\right),u v \rho ,\frac{1}{2} \rho  \left(2c_v T (2 \gamma -1)+u^2+3 v^2\right),c_v v (2 \gamma -1) \rho \right\}
%\end{align*}
%The viscous Jacobians become (when linearized with respect to $\{\sigma_{11},\sigma_{12}, \sigma_{22}\}$)
%\begin{align*}
%G^1_{2,\boldsymbol \sigma} &= \{1,0,0\}\\
%G^2_{2,\boldsymbol \sigma} &= \{0,1,0\}\\
%G^1_{3,\boldsymbol \sigma} &= \{0,1,0\}\\
%G^2_{3,\boldsymbol \sigma} &= \{0,0,1\}\\
%G^1_{4,\boldsymbol \sigma} &= \{u_1,u_2,0\}\\
%G^2_{4,\boldsymbol \sigma} &= \{0,u_1,u_2\}
%\end{align*}
%and (when linearized with respect to the Eulerian variables)
%\begin{align*}
%G^1_{4,\boldsymbol U} &= \{0,\sigma_{11},\sigma_{12},0\}\\
%G^2_{4,\boldsymbol U} &= \{0,\sigma_{12},\sigma_{22},0\}
%\end{align*}

@@

Recall the convection-diffusion problem 
\begin{align*}
\div \left(\beta u - \sigma\right) &= f\\
\frac{1}{\epsilon}\sigma - \grad u &= 0.
\end{align*}
On domain $\Omega$, with mesh $\Oh$ and mesh skeleton $\Gh$, the DPG variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh} - \LRa{
\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{ \widehat{f}_n,
  \jump{v} }_{\Gh}.
\end{align*}
with $v\in H^1$ and $\tau \in H({\rm div},\Oh)$. The test norm adopted for convection-diffusion in Section~\secref{sec:testNormSec} and in \cite{DPGrobustness2} is defined elementwise on $K$ as
\[
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2.
\]
This test norm both delivers robust control of the error in the $L^2$ variables $u$ and $\sigma$ and avoids boundary layers in the computation of local test functions.

Our test norm is extrapolated to the Navier-Stokes equations as follows: we denote the vector of $H^1$ test functions as $v=\{v_1,v_2,v_3,v_4\}$, and similarly for $W = \{\tau_1,\tau_2,\tau_3\}$. Similarly, we group our Eulerian and stress variables into the vector variables $U$ and $\Sigma$, respectively. If $R_{\rm Euler}(U,\Sigma)$ and $R_{\rm visc}(U,\Sigma)$ are Eulerian and viscous nonlinear residuals, our formulation for the linearized Navier-Stokes equations can be written as
\begin{align*}
\div \left(A_{\rm Euler}U - A_{\rm visc}\Sigma\right) &= R_{\rm Euler}(U,\Sigma)\\
E_{\rm visc} \Sigma - \grad U &= R_{\rm visc}(U,\Sigma)
\end{align*}
with variational formulation
\begin{align*}
\LRa{\widehat{F}_n,V}_{\Gh} + \left(U,\div W - A_{\rm Euler}^T\grad  V\right) + \LRa{\widehat{U},W\cdot n}_{\Gh} + \left(\Sigma,E_{\rm visc}^TW - A_{\rm visc}^T\grad  V\right) &= 0
\end{align*}
Identifying vector-valued terms in the Navier-Stokes formulation with equivalent scalar terms in the convection-diffusion equation allows us to extrapolate our test norm to systems of equations
\begin{align*}
\|\left(V,W\right)\|_{V,K}^2 =& \min\left\{\frac{1}{\Reyn|K|},1\right\}\|v\|^2 + \frac{1}{\Reyn} \|A_{\rm visc}^T\grad v\|^2 + \|A_{\rm Euler}^T \grad v\|^2 \\
& + \| \div W\|^2 + \min\left\{\Reyn,\frac{1}{|K|}\right\}\|E_{\rm visc}^T\tau\|^2.
\end{align*}
An advantage of this extrapolation approach is that the incompletely parabolic nature of the Navier-Stokes equation is taken into account; there is no diffusive term present in the mass conservation equation, and the test norm reflects that by requesting only limited regularity of $v_1$.\footnote{The situation is analogous to using the full $H^1(\Oh)$ norm for the pure convection equation --- the optimal test norm $\nor{v}_V = \nor{\beta\cdot \grad v} + \nor{v}$ implies only streamline regularity, whereas taking $\nor{v}_V = \nor{\grad v} + \nor{v}$ implies stronger regularity on the test space $V$ than the graph norm. Consequently, convergence is suboptimal for DPG applied to the convection problem under the $H^1(\Oh)$ test norm.}

@@

>>>>

We briefly review the compressible Navier-Stokes equations, given in Section~\secref{sec:compNS}, and formulate DPG for the nonlinear system. 
\begin{itemize}
\item \textbf{Conservation equations}
\begin{align*}
\div \vecttwo{\rho u_1 }{\rho u_2} &= 0\\
\div \vecttwo{\rho u_1^2+p }{\rho u_1 u_2} &= \div\left(\vec{\sigma_{i1}}\right)\\
\div \vecttwo{\rho u_1 u_2}{\rho u_2^2+p } &= \div\left(\vec{\sigma_{i2}}\right)\\
\div\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} &= \div\left[\boldsymbol\sigma \boldsymbol u + \vec{q}\right],
\end{align*}
where $\boldsymbol \sigma$ is the stress tensor whose $ij$th term is $\sigma_{ij}$, and $\boldsymbol u$ is the vector $(u_1,u_2)^T$.

@@

We represent $\boldsymbol\sigma$ using the Newtonian fluid law
\[
\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}
\]
where $\mu$ is viscosity and $\lambda$ is bulk viscosity. 
%Following elasticity, a general formula for the stress tensor is 
%\[
%\sigma_{ij} = 2\mu \epsilon_{ij} + \lambda \epsilon_{kk} \delta_{ij}
%\]
We can invert the stress tensor under isotropic and plane strain assumptions to get
\[
\frac{1}{2}\left(\grad  \boldsymbol u + \grad ^T  \boldsymbol u\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
We also have
\[
\frac{1}{2}\left(\grad  \boldsymbol u + \grad ^T  \boldsymbol u\right) = \grad  \boldsymbol u - \boldsymbol \omega
\]
where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
\[
\boldsymbol \omega = \frac{1}{2}\left(\grad  \boldsymbol u - \grad ^T  \boldsymbol u\right).
\]
Thus our final form is
\begin{align*}
\grad  \boldsymbol u - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
Notice that $\boldsymbol \omega$ is implicitly defined to be the antisymmetric part of $\grad u$ by taking the symmetric part of the above equation.

@@

To summarize, our system of equations in the classical variables is now
\begin{align*}
\div \vecttwo{\rho u_1 }{\rho u_2} &= 0\\
\div \left(\vecttwo{\rho u_1^2+p }{\rho u_1 u_2} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u_1 u_2}{\rho u_2^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} - \boldsymbol\sigma \mathbf{u} - \vec{q}\right) &=0\\
\frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I &= \grad \mathbf{u} - \Reyn \, {\boldsymbol \omega}\\
\frac{1}{\kappa}\vec{q} &= \grad T
\end{align*}
We strongly enforce symmetry of the stress tensor $\boldsymbol \sigma$ by setting $\sigma_{21} = \sigma_{12}$. Additionally, we have scaled the antisymmetric tensor $\boldsymbol \omega$ by the Reynolds number to ensure that $\boldsymbol \omega = O(1)$ for all ranges of $\Reyn$.

@@

As the equations for viscous compressible flow are nonlinear and cannot be solved exactly, we linearize the equations and adopt an iterative procedure for approximating the nonlinear solution.\footnote{We note that it is possible to linearize the strong form of the equations and then derive a linearized weak form, instead of linearizing the weak form of the nonlinear equations, which is done here.  The two formulations are equivalent; however, extraneous linearization of the fluxes is avoided using the latter approach.}  We outline the linearization of both the conservation and stress laws in this section.

@@

The Navier-Stokes conservation laws can be written as 
\begin{align*}
\div \vecttwo{\rho u_1 }{\rho u_2} &= 0\\
\div \left(\vecttwo{\rho u_1^2+p }{\rho u_1 u_2} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u_1 u_2}{\rho u_2^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u - \vec{q}\right) &=0,
\end{align*}
where $\boldsymbol \sigma_1$ is the $i$th column or row of $\boldsymbol \sigma$, or more generally, if we group our Eulerian and stress variables into the vector variables $\boldsymbol U$ and $\boldsymbol \Sigma$, respectively
\[
\div (F_i(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \Sigma) = 0, \qquad i = 1,\ldots, 4,
\]
where $F_i$and $G_i$ are given as
\begin{align*}
F_1 = \vecttwo{\rho u_1 }{\rho u_2}, \quad &G_1 = 0\\
F_2 = \vecttwo{\rho u_1^2+p }{\rho u_1 u_2},\quad &G_2 = \boldsymbol \sigma_{1}\\
F_3 = \vecttwo{\rho u_1 u_2}{\rho u_2^2+p }, \quad &G_3 = \boldsymbol \sigma_{2}\\
F_4 = \vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2}, \quad &G_4 = \boldsymbol \sigma_1 \cdot \boldsymbol u + \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}
\end{align*}
The variational form restricted to a single element gives
\[
\langle \widehat{F}_i\cdot n, v\rangle - \int_K  (F(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \Sigma)) \cdot \grad v_i = 0 , \qquad i = 1,\ldots, 4
\]
and the variational form over the entire domain is given by summing up the element-wise contributions.

The presence of terms such as $\boldsymbol \sigma_i \cdot \boldsymbol u$ means that we will need to linearize in the stress variables $\sigma_ij$ in addition to our Eulerian quantities. Since fluxes and traces are linear functions of the unknowns, we do not need to linearize them. Instead, fluxes $\widehat{F}_{i,n}$ and traces $\widehat{u},\widehat{v},\widehat{T}$ will represent normal traces and traces of the accumulated nonlinear solution. The linearized variational formulation is thus
\begin{align*}
\langle \widehat{F}_i\cdot n, v\rangle &- \int_K  \left(F_{i,\boldsymbol U}(\boldsymbol U)\cdot \Delta \boldsymbol U -G_{i,\boldsymbol U}(\boldsymbol U, \boldsymbol \Sigma)\cdot \Delta \boldsymbol U - G_{i,\boldsymbol \Sigma}(\boldsymbol U, \boldsymbol \Sigma)\cdot \Delta \boldsymbol \Sigma \right)\cdot \grad v_i \\
&= \int_K  \left(F_i(\boldsymbol U)-G_i(\boldsymbol U)\right) \cdot \grad v_i \\
\qquad i &= 1,\ldots, 4
\end{align*}
where $F^i_{j,\boldsymbol U}$, $G^i_{j,\boldsymbol U}$, and $G^i_{j,\boldsymbol \Sigma}$ are the Eulerian and two viscous Jacobians (linearized w.r.t.\ the Eulerian/viscous variables), respectively.
%where
%\begin{align*}
%F^1_{1,\boldsymbol U} &= \{u,\rho ,0,0\} \\
%F^2_{1,\boldsymbol U} &= \{v,0,\rho ,0\} \\
%F^1_{2,\boldsymbol U} &=\left\{c_v T (\gamma -1)+u^2,2 u \rho ,0,c_v (\gamma -1) \rho \right\}\\
%F^2_{2,\boldsymbol U} &=\{u v,v \rho ,u \rho ,0\}\\
%F^1_{3,\boldsymbol U} &=\{u v,v \rho ,u \rho ,0\}\\
%F^2_{3,\boldsymbol U} &=\left\{c_v T (\gamma -1)+v^2,0,2 v \rho ,c_v (\gamma -1) \rho \right\}\\
%F^1_{4,\boldsymbol U} &=\left\{\frac{1}{2} u \left(2 c_v T (2 \gamma -1)+u^2+v^2\right),\frac{1}{2} \rho  \left(2 c_v T (2 \gamma -1)+3 u^2+v^2\right),u v \rho ,c_v u (2 \gamma -1) \rho \right\}\\
%F^2_{4,\boldsymbol U} &=\left\{\frac{1}{2} v \left(2 c_v T (2 \gamma -1)+u^2+v^2\right),u v \rho ,\frac{1}{2} \rho  \left(2c_v T (2 \gamma -1)+u^2+3 v^2\right),c_v v (2 \gamma -1) \rho \right\}
%\end{align*}
%The viscous Jacobians become (when linearized with respect to $\{\sigma_{11},\sigma_{12}, \sigma_{22}\}$)
%\begin{align*}
%G^1_{2,\boldsymbol \sigma} &= \{1,0,0\}\\
%G^2_{2,\boldsymbol \sigma} &= \{0,1,0\}\\
%G^1_{3,\boldsymbol \sigma} &= \{0,1,0\}\\
%G^2_{3,\boldsymbol \sigma} &= \{0,0,1\}\\
%G^1_{4,\boldsymbol \sigma} &= \{u_1,u_2,0\}\\
%G^2_{4,\boldsymbol \sigma} &= \{0,u_1,u_2\}
%\end{align*}
%and (when linearized with respect to the Eulerian variables)
%\begin{align*}
%G^1_{4,\boldsymbol U} &= \{0,\sigma_{11},\sigma_{12},0\}\\
%G^2_{4,\boldsymbol U} &= \{0,\sigma_{12},\sigma_{22},0\}
%\end{align*}

@@

Recall the convection-diffusion problem 
\begin{align*}
\div \left(\beta u - \sigma\right) &= f\\
\frac{1}{\epsilon}\sigma - \grad u &= 0.
\end{align*}
On domain $\Omega$, with mesh $\Oh$ and mesh skeleton $\Gh$, the DPG variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh} - \LRa{
\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{ \widehat{f}_n,
  \jump{v} }_{\Gh}.
\end{align*}
with $v\in H^1$ and $\tau \in H({\rm div},\Oh)$. The test norm adopted for convection-diffusion in Section~\secref{sec:testNormSec} and in \cite{DPGrobustness2} is defined elementwise on $K$ as
\[
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau-\beta\cdot\grad v\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2.
\]
This test norm both delivers robust control of the error in the $L^2$ variables $u$ and $\sigma$ and avoids boundary layers in the computation of local test functions.

This test norm is extrapolated to the Navier-Stokes equations as follows: we denote the vector of $H^1$ test functions as $\boldsymbol V=\{v_1,v_2,v_3,v_4\}$, and similarly for $\boldsymbol W = \{\tau_1,\tau_2,\tau_3\}$. If $R_{\rm Euler}(\boldsymbol U,\boldsymbol \Sigma)$ and $R_{\rm visc}(\boldsymbol U,\boldsymbol \Sigma)$ are Eulerian and viscous nonlinear residuals, our formulation for the linearized Navier-Stokes equations can be written as
\begin{align*}
\div \left(A_{\rm Euler}\boldsymbol \delta U - A_{\rm visc}\boldsymbol \delta \Sigma\right) &= R_{\rm Euler}(\boldsymbol U,\boldsymbol \Sigma)\\
E_{\rm visc}\boldsymbol  \delta \Sigma - \grad \boldsymbol \delta U &= R_{\rm visc}(\boldsymbol U,\boldsymbol \Sigma)
\end{align*}
with variational formulation
\begin{align*}
\LRa{\widehat{F}_n,\boldsymbol V}_{\Gh} + \left(\boldsymbol \delta U,\div \boldsymbol W - A_{\rm Euler}^T\grad \boldsymbol  V\right) + \LRa{\widehat{\boldsymbol U},\boldsymbol W\cdot n}_{\Gh} + \left(\boldsymbol \delta \Sigma,E_{\rm visc}^T\boldsymbol W - A_{\rm visc}^T\grad \boldsymbol  V\right) &= 
\\ \LRa{R_{\rm Euler}(\boldsymbol U,\boldsymbol \Sigma),\boldsymbol V} + \LRa{R_{\rm visc}(\boldsymbol U,\boldsymbol \Sigma),\boldsymbol W}
\end{align*}
Identifying vector-valued terms in the Navier-Stokes formulation with equivalent scalar terms in the convection-diffusion equation allows us to extrapolate our test norm to systems of equations
\begin{align*}
\|\left(\boldsymbol V,\boldsymbol W\right)\|_{V, K}^2 =& \|\boldsymbol V\|^2 + \frac{1}{\Reyn} \|A_{\rm visc}^T\grad \boldsymbol V\|^2 + \|A_{\rm Euler}^T \grad \boldsymbol V\|^2 \\
& + \| \div \boldsymbol W - A_{\rm Euler}^T \grad \boldsymbol V\|^2 + \min\left\{\Reyn,\frac{1}{|K|}\right\}\|E_{\rm visc}^T\boldsymbol W\|^2.
\end{align*}
An advantage of this extrapolation approach is that the incompletely parabolic nature of the Navier-Stokes equation is taken into account; there is no diffusive term present in the mass conservation equation, and the test norm reflects that by requesting only limited regularity of $v_1$, the test function for the conservation equation.\footnote{The situation is analogous to using the full $H^1(\Oh)$ norm for the pure convection equation --- the optimal test norm $\nor{v}_V = \nor{\beta\cdot \grad v} + \nor{v}$ implies only streamline regularity, whereas taking $\nor{v}_V = \nor{\grad v} + \nor{v}$ implies stronger regularity on the test space $V$ than the graph norm. Consequently, convergence is suboptimal for DPG applied to the convection problem under the $H^1(\Oh)$ test norm.}

@@

\section{Nonlinear solver}

For our nonlinear solver, we use a pseudo-time stepping approach to iterate to a steady state solution, along with a greedy refinement scheme to eliminate spatial discretization error.  We cover briefly the details of the pseudo-timestepping method in this section.  
%Though we have not yet implemented adaptive timestepping, we are able to take large enough uniform timesteps over a coarse mesh such that convergence occurs sufficiently quickly.

\subsection{Pseudo-timestepping}

The solution of the compressible Navier-Stokes equations can be quite challenging; as mentioned previously, the direct application of a Newton algorithm often will not converge, especially for high Reynolds numbers.  Typically, a pseudo-timestepping algorithm is used in lieu of a full nonlinear Newton algorithm. The pseudo-timestep proceeds as follows: given the transient terms present in the conservation laws of compressible flow
\begin{align*}
\pd{\rho}{t} + \ldots \\
\pd{\LRp{\rho u}}{t} + \ldots \\
\pd{\LRp{\rho v}}{t} + \ldots \\
\pd{\LRp{\rho e}}{t} + \ldots,
\end{align*}
we discretize each time derivative using an implicit timestepping method.  Though second order time discretizations have been shown to be effective \cite{Demkowicz1990275, BenKirk}, we choose a first order backwards Euler discretization for simplicity.  Due to the fact that we've chosen to solve the Navier-Stokes equations under primitive variables $\rho, u, v$, and $T$, the time terms are nonlinear as well, and must be linearized.  After time discretization and linearization, we are left with a coupled system of reaction terms in the problem for our solution update at every timestep
\begin{align*}
A_{\rm time}\boldsymbol \delta U_i + \div \left(A_{\rm Euler}\boldsymbol \delta U_i - A_{\rm visc}\boldsymbol \delta \Sigma_i\right) &= R_{\rm Euler}(\boldsymbol U_i,\boldsymbol \Sigma_i) + R_{\rm time}(\boldsymbol U_{i-1},\boldsymbol U_i)\\
E_{\rm visc} \boldsymbol \Sigma_i - \grad \boldsymbol U_i &= R_{\rm visc}(\boldsymbol U_i,\boldsymbol \Sigma_i),
\end{align*}
where $R_{\rm time}(\boldsymbol U_{i-1},\boldsymbol U_i)$ is the transient residual.  Including the transient terms in our test norm as well, our final test norm is 
\begin{align*}
\|\left(\boldsymbol V,\boldsymbol W\right)\|_{V,K}^2 =& \nor{A_{\rm time}^T\boldsymbol V} + \|\boldsymbol V\|^2 + \frac{1}{\Reyn} \|A_{\rm visc}^T\grad \boldsymbol V\|^2 + \|A_{\rm Euler}^T \grad \boldsymbol V\|^2 \\
& + \| \div \boldsymbol W - A_{\rm Euler}^T \grad \boldsymbol V\|^2 + \min\left\{\Reyn,\frac{1}{|K|}\right\}\|E_{\rm visc}^T\boldsymbol W\|^2.
\end{align*}
Finally, convergence of the pseudo-timestepping method is determined by measuring the transient residual in the energy norm; in other words, we measure $\nor{e_{\rm time}}_V$, where
\[
\LRp{\LRp{\boldsymbol V,\boldsymbol W},\LRp{\boldsymbol dV,\boldsymbol dW}}_V = \LRp{R_{\rm time}(\boldsymbol U_{i-1},\boldsymbol U_i),\boldsymbol V}_{\L}, \quad \forall \LRp{\boldsymbol dV,\boldsymbol dW} \in V.
\]
The energy norm thus provides a consistent measure in which convergence of the nonlinear iteration at each timestep, convergence of the pseudo-timestepping algorithm to steady state, and nonlinear residual can all be assessed.

\subsubsection{Dependence of solution on $dt$}

A surprising feature of pseudo-timestepping schemes for DPG is that, under the problem-dependent minimum-residual nature of the method, convergence to steady state can yield qualitatively slightly different solutions under different size timesteps.  We illustrate this using the plate example for convection-diffusion, which we described first in Section~\secref{sec:confusionPlate}.  We consider the transient form of the conservation equation for convection-diffusion
\[
\pd{u}{t} + \div \LRp{\beta u - \sigma} = f, 
\]
where the stress law remains unchanged.  Applying a backwards Euler time discretization, at each timestep, we will solve for the current solution $u_i$ (as well as the current stress $\sigma_i$) given the previous timestep solution $u_{i-1}$ under
\[
\frac{u_i}{dt} + \div \LRp{\beta u_i - \sigma_i} = f + \frac{u_{i-1}}{dt}.
\]
The conserved flux $\beta_n u - \sigma_n$ is set to freestream values on the inflow, and 0 on the top boundary $y = 1$ and half of the bottom boundary $x < .5, y = 0$.  We set $u = 1$ for the boundary $x \in (.5,1), y = 0$. For this example, $\beta = (1,0)$ and $\epsilon = 10^{-3}$.  
\begin{figure}
\centering
\subfigure[$dt = .1$]{\includegraphics[scale=.35]{figs/confusionDtFigs/dt10.png}}
\subfigure[$dt = .01$]{\includegraphics[scale=.35]{figs/confusionDtFigs/dt100.png}}
\caption{Comparison of pseudo-timestepping to steady state for a convection-diffusion problem under two different sizes of timestep.}
\label{fig:dtComparison}
\end{figure}
This can be understood as the consequence of solving a ``moving target'' optimization problem -- our variational formulation and test norm for this problem are
\begin{align*}
\LRp{u_{i}, \frac{1}{dt}v}_{\L} + b\LRp{u_i,v} &= l(v) + \LRp{u_{i-1},\frac{1}{dt}v}_{\L}\\
\nor{\LRp{\tau,v}}_{V,dt}^2 &= \frac{1}{dt}\nor{v}_{\L}^2 + \nor{v}_V^2,
\end{align*}
where $\nor{\LRp{\tau,v}}_{V}$ is the coupled test norm introduced in Section~\secref{sec:testNormSec}, and $b(u,v)$ and $l(v)$ are the bilinear form and load for the steady state form of the convection diffusion equation.  DPG solutions minimize the functional 
\[
J(u_i) = \sup_{v\in V} \frac{\LRp{u_{i}-u_{i-1}, \frac{1}{dt}v}_{\L} + b(u_i,v) - l(v)}{\LRp{\frac{1}{dt}\nor{v}_{\L}^2 + \nor{v}_V^2}^{\frac{1}{2}}}
\]
over a given mesh.  As $u_{i-1}\rightarrow u_{i}$, which we expect to happen as the pseudo-timestepping algorithm converges to steady state, the minimized functional becomes
\[
J(u_i) = \sup_{v\in V} \frac{b(u_i,v) - l(v)}{\LRp{\frac{1}{dt}\nor{v}_{\L}^2 + \nor{v}_V^2}^{\frac{1}{2}}}.
\]
While the transient portion of the residual disappears, a factor of $\frac{1}{dt}$ is still present in the test norm.\footnote{While the solution under smaller $dt$ appears to give visually higher quality results, we stress that simply adding the term $\frac{1}{dt}\nor{v}_{\L}$ to the test norm under the steady state version of the convection-diffusion equations does not achieve the same effect.  We are able to add this term without negative consequence due to the inclusion of the $\frac{u_i}{dt}$ term present in the variational formulation (see Chapter~\secref{chap:robustDPG} and Appendix~\secref{appendix:timesteppingConfusion} for mathematical details).  Numerical experiments indicate that including $\alpha\nor{v}_{\L}$ in the test norm, where $\alpha > \frac{1}{dt}$, converges much more slowly to a fine-mesh reference solution than if $\alpha = \frac{1}{dt}$.}  Thus, we can expect that the nature of the steady state solution achieved through convergence of the pseudo-time algorithm can depend on the timestep $dt$.  We observe the same phenomena for analogous problems in compressible flow as well.

\subsubsection{Adaptive time thresholding}
